Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-09-14 paper/current/main.tex
add data part

%
% Figure - Gravity law performance
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/GravityLawFig/gravity_performance.pdf}
	\caption{
		\textbf{Embedding distance encodes functional distance and better predicts mobility.}
		\textbf{a-b.} Embedding distance (cosine distance between organization vectors) better explains the expected flux between organization pairs than geographic distance.
		The red line is the line of the best fit.
		Black dots are mean flux across binned distances.
		99\% confidence intervals are plotted, but are too small to be visible.
		Correlation is calculated on the data in the $\log-\log$ scale. 
		The lightness of each hex bin indicates the frequency of organization pairs within it.
		\textbf{c-d.} Predictions of flux between organization pairs made using embedding distance outperform those made using geographic distance.
		Boxplots show distribution of actual flux for binned values of predicted flux.
		Box color corresponds to the degree to which the distribution overlaps with $y = x$;
		a perfect prediction yields all points on the black line.
		``RMSE'' is the root-mean-squared error between the actual and predicted values.
		\textbf{e-f.} Predictions made using model parameters learned from a subset of domestic (within-country) organization pairs.
		Embedding distance outperforms those made for geographic distance.
		\textbf{g-h.} Embedding distance also outperforms geographic distance when predicting flux using model parameters estimated from only international mobility (across-country organization pairs).
		Predictions of domestic and international mobility are qualitatively similar whether predicted using models with only within- or across-country pairs (\textbf{e-h}), or all organizations pairs (Fig. \ref{fig:supp:gravity_prediction_subsets}).
	}
	\label{fig:gravity_performance}
\end{figure}

@@

\subsection*{Data}
%% Main dataset
% Do we need to talk about how we define mobility in relation to past papers?
We source co-affiliation trajectories of authors from the Web of Science database hosted by the Center for Science and Technology Studies at Leiden University.
Trajectories are constructed from author affiliations listed on the byline of publications for a given author.
Given the limitations of author-name disambiguation, we limit our analyses to papers published after 2008, when the Web of Science began providing full names and institutional affiliations~\autocite{caron2014disambiguation} that improved disambiguation (see \SI).
This yields 33,934,672 author-affiliation combinations representing 12,963,792 authors.
Each author-affiliation combination is associated with the publication year and an ID linking it to one of 8,661 disambiguated organizational affiliations (see \SI for more detail).
Trajectories are represented as the list of author-affiliation combinations, ordered by year of publication, and randomly ordered for combinations within the same year. The most fine-grained geographical unit in this data is the organization, such as a ``university'', ``research institute'', ``business'', or ``government agency''.

@@

>>>>

%
% Figure - Gravity law performance
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/GravityLawFig/gravity_performance.pdf}
	\caption{
		\textbf{Embedding distance encodes functional distance and better predicts mobility.}
		\textbf{a-b.} Embedding distance (cosine distance between organization vectors) better explains the expected flux between organization pairs than geographic distance.
		The red line is the line of the best fit.
		Black dots are mean flux across binned distances.
		99\% confidence intervals are plotted, but are too small to be visible.
		Correlation is calculated on the data in the log-log scale. 
		The lightness of each hex bin indicates the frequency of organization pairs within it.
		\textbf{c-d.} Predictions of flux between organization pairs made using embedding distance outperform those made using geographic distance.
		Boxplots show distribution of actual flux for binned values of predicted flux.
		Box color corresponds to the degree to which the distribution overlaps with $y = x$;
		a perfect prediction yields all points on the black line.
		``RMSE'' is the root-mean-squared error between the actual and predicted values.
		\textbf{e-f.} Predictions made using model parameters learned from a subset of domestic (within-country) organization pairs.
		Embedding distance outperforms those made for geographic distance.
		\textbf{g-h.} Embedding distance also outperforms geographic distance when predicting flux using model parameters estimated from only international mobility (across-country organization pairs).
		Predictions of domestic and international mobility are qualitatively similar whether predicted using models with only within- or across-country pairs (\textbf{e-h}), or all organizations pairs (Fig. \ref{fig:supp:gravity_prediction_subsets}).
	}
	\label{fig:gravity_performance}
\end{figure}

@@

\subsection*{Data - Scientific mobility}
%% Main dataset
% Do we need to talk about how we define mobility in relation to past papers?
We source co-affiliation trajectories of authors from the Web of Science database hosted by the Center for Science and Technology Studies at Leiden University.
Trajectories are constructed from author affiliations listed on the byline of publications for a given author.
Given the limitations of author-name disambiguation, we limit our analyses to papers published after 2008, when the Web of Science began providing full names and institutional affiliations~\autocite{caron2014disambiguation} that improved disambiguation (see \SI).
This yields 33,934,672 author-affiliation combinations representing 12,963,792 authors.
Each author-affiliation combination is associated with the publication year and an ID linking it to one of 8,661 disambiguated organizational affiliations (see \SI for more detail).
Trajectories are represented as the list of author-affiliation combinations, ordered by year of publication, and randomly ordered for combinations within the same year. The most fine-grained geographical unit in this data is the organization, such as a ``university'', ``research institute'', ``business'', or ``government agency''.

@@

\subsection*{Data -US Airport Itinerary data}
We use US airport itinerary data from the Bureau of Transportation statistics provided by the United States Department of Transportation. Origin and Destination Survey (DB1B) is a 10\% sample of domestic airline tickets, including origin, destination, and other itinerary details. Trajectories are represented as the list of airport code, cover itinerary between 1993 to 2020 and 30,776,0841 trajectory in total and the unique number of airports is 838.
